\chapter{Quality Assurance: Verification and Validation}
\label{ch:quality}

This chapter documents the quality-assurance programme of the Smart Home Energy
Optimizer (SHEO). It establishes the conceptual vocabulary of software quality,
positions the project's testing effort within the classical V-model, and then
presents the testing campaign \emph{organised by V-model level} --- unit,
integration and system testing --- as required for the client acceptance test.
Each level opens with a statement of its objectives, scope and approach, and then
lists its test cases in a uniform $\langle$test case, input, expected output,
status$\rangle$ form, with the design technique (a standard white-box or black-box
technique) indicated for each. The discussion is grounded in the project's
\textbf{67 automated tests} and concludes with a requirements-to-tests
traceability summary and a forward-looking classification of anticipated
maintenance.

\section{Software Quality, Verification and Validation}
\label{sec:qa-vv}

Following the IEEE Standard Glossary of Software Engineering Terminology
(IEEE~610.12)~\cite{ieee610}, \emph{software quality} is the degree to which a system meets
specified requirements and, in turn, the degree to which it meets stated and
implied user needs and expectations. Quality is therefore not a single attribute
but a composite of functional correctness together with non-functional
properties such as reliability, security, safety and maintainability. For SHEO
the load-bearing non-functional families are safety (NFR-SAF, protecting
safety-critical appliances and rate-limiting the air-conditioner compressor) and
security (NFR-SEC, authentication, role-based access control and never storing
plaintext secrets).

Two complementary activities establish quality. \emph{Verification} asks
``are we building the product right?'' --- it is the developer-facing check that
each artefact conforms to its specification and to the artefact that preceded it
(requirements, design, code). \emph{Validation} asks ``are we building the right
product?'' --- it is the user-facing check that the delivered system actually
satisfies the stakeholder's real needs. In SHEO, verification is realised by the
unit and integration tests that confirm the implementation honours its
contracts, while validation is realised by the system tests that
exercise the application through its public interface against the acceptance
criteria agreed with the customer --- here the SHEO stakeholder in the IEEE
sense, namely the Administrator who commissions and accepts the system. The two
are not interchangeable: a system can
be verified (faithful to its specification) yet invalid (the specification was
wrong), which is precisely why both arms of the V-model below are required.

\section{The V-model}
\label{sec:qa-vmodel}

The \emph{V-model} arranges the software life-cycle so that every constructive
(development) phase on the descending left arm is paired with a corresponding
destructive (test) phase on the ascending right arm: requirements analysis with
acceptance testing, architectural design with system testing, detailed design with
integration testing, and coding with unit testing. The test cases for a level are
derived from, and validate against, the artefact produced by the phase opposite it ---
acceptance tests against the requirements, system tests against the architecture,
integration tests against the detailed design, and so on. Crucially, the model also
dictates \emph{which family of test-case design technique} dominates at each level:
\textbf{white-box} (structural) techniques are applied where the internal structure
is the object under test --- unit and integration testing --- while \textbf{black-box}
(specification-based) techniques are applied where only externally visible behaviour
matters --- system and acceptance testing.

SHEO populates every level of the right arm. \emph{Unit testing} exercises the
pure domain functions in isolation; \emph{integration testing} drives the wired
stack (API router $\rightarrow$ service $\rightarrow$ repository $\rightarrow$
persistence) across real boundaries; and \emph{system testing} asserts aggregate,
user-visible outputs end-to-end on a fully seeded application. \emph{Acceptance
testing} is then realised by re-running the system suite against the demonstration
seed in front of the customer, validating the acceptance focus items.
The sequence-level interactions that integration tests confirm are illustrated in
the design chapter (Figure~\ref{fig:seq-control}).

\section{Test Strategy and Levels}
\label{sec:qa-strategy}

The automated suite comprises \textbf{67 tests}. It runs in-process against the
application's public interface using an in-memory test harness, requiring no live
server, network or hardware; a disposable database is recreated for each run, so
the suite is fully deterministic. The split between levels is physical: pure-logic
unit tests touch no transport layer and no database, whereas every other suite
boots a fully seeded application and drives it through its public endpoints.

Each test is expressed in the canonical test-case form
$\langle$input, expected outcome$\rangle$: an input (or input sequence) is applied
and the actual result is compared against an \emph{expected output that is defined
in advance}, never inferred after observing the program. A collection of such
cases that share a fixture and a target constitutes a \emph{test suite} --- the
project organises its tests into nine such suites. The strategy observes three
classical testing principles:

\begin{itemize}[leftmargin=1.4em]
  \item the \textbf{tester $\ne$ developer} separation of concern: tests are
        written from the specification (the requirements and the capability
        contract) rather than to ratify whatever the code happens to do;
  \item the \textbf{expected output is fixed in advance}, so a test can only pass
        by the implementation being correct, not by redefining correctness;
  \item \textbf{regression}: because the suite is automated and deterministic, it
        is re-run on every change to guard previously verified behaviour against
        accidental breakage.
\end{itemize}

Table~\ref{tab:levels} summarises how the levels, their objectives, the dominant
design-technique family, and the suites that populate them line up. The three
following sections take each level in turn; in their test-case tables the
\emph{Status} column records the outcome of the most recent run --- all
\textbf{67 tests pass}.

\begin{table}[htbp]
\centering
\caption{Test levels, their dominant test-case design technique, and the covering
suites.}
\label{tab:levels}
\begin{tabularx}{\textwidth}{@{}l L{0.30\textwidth} L{0.24\textwidth} Y@{}}
\toprule
\textbf{Level} & \textbf{Objective / scope} & \textbf{Technique family} & \textbf{Suites} \\
\midrule
Unit        & Pure domain functions in isolation (no HTTP, no DB) &
              White-box: statement, branch and path coverage &
              \texttt{unit\_domain} \\
Integration & Wired components cooperating across real boundaries
              (router/service/repository/persistence, event bus, adapters) &
              White-box: branch and decision coverage of the integrated path &
              \texttt{devices\_capability}, \texttt{rules}, \texttt{safety} \\
System      & End-to-end, user-visible behaviour through the public REST API &
              Black-box: equivalence partitioning, boundary-value analysis,
              decision tables, state-based &
              \texttt{monitoring}, \texttt{savings}, \texttt{recommendations},
              \texttt{auth\_rbac}, \texttt{admin} \\
\bottomrule
\end{tabularx}
\end{table}

\section{Unit Testing (White-box)}
\label{sec:qa-unit}

\paragraph{Objective.} Verify the algorithms and internal structure of the
project's pure domain functions in isolation, confirming each computes the right
result on every branch of its logic.

\paragraph{Scope.} The functions with no transport or persistence dependency: the
capability \emph{validator}, the tariff math (flat and tiered pricing), the
billing-cycle and time-window utilities, the simulator energy model, and password
hashing. These are reached by direct function call --- no HTTP, no database.

\paragraph{Approach.} \emph{Control-flow testing}: for each function we reason
about its control flow graph (CFG) and select concrete inputs that achieve
\textbf{statement}, \textbf{branch} and, for the small functions, full
\textbf{path} coverage. As the worked example, take the air-conditioner
target-temperature guard, whose valid range is the fixed, deterministic interval
$16$--$30\,$\textdegree C. It rejects a non-numeric value, rejects a value below
$16$, rejects a value above $30$, and otherwise accepts. Its CFG
(Figure~\ref{fig:cfg}) has three decisions --- nodes 2, 4 and 6 --- and four
feasible paths.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    node distance=8mm and 16mm,
    block/.style={draw, rounded corners, align=center, font=\scriptsize,
                  minimum width=20mm, minimum height=7mm},
    dec/.style={draw, diamond, aspect=2, align=center, font=\scriptsize,
                inner sep=1pt},
    term/.style={draw, ellipse, align=center, font=\scriptsize, fill=red!8},
    ok/.style={draw, ellipse, align=center, font=\scriptsize, fill=green!12},
    edge/.style={-{Latex[length=2mm]}}
]
\node[block] (n1) {1: read $v$};
\node[dec]   (n2) [below=of n1] {2: numeric?};
\node[term]  (n3) [right=of n2] {3: reject\\``not numeric''};
\node[dec]   (n4) [below=of n2] {4: $v < 16$?};
\node[term]  (n5) [right=of n4] {5: reject\\``below 16''};
\node[dec]   (n6) [below=of n4] {6: $v > 30$?};
\node[term]  (n7) [right=of n6] {7: reject\\``above 30''};
\node[ok]    (n8) [below=of n6] {8: accept $v$};

\draw[edge] (n1) -- (n2);
\draw[edge] (n2) -- node[above,font=\tiny]{no}  (n3);
\draw[edge] (n2) -- node[right,font=\tiny]{yes} (n4);
\draw[edge] (n4) -- node[above,font=\tiny]{yes} (n5);
\draw[edge] (n4) -- node[right,font=\tiny]{no}  (n6);
\draw[edge] (n6) -- node[above,font=\tiny]{yes} (n7);
\draw[edge] (n6) -- node[right,font=\tiny]{no}  (n8);
\end{tikzpicture}
\caption{Control flow graph of the air-conditioner target guard, with the fixed
deterministic range $16$--$30\,$\textdegree C. Diamonds are decisions; red
terminals are rejections; the green terminal accepts. The three decisions yield
four feasible paths, each reachable by a concrete input.}
\label{fig:cfg}
\end{figure}

Each path corresponds to one concrete input. Table~\ref{tab:cfg-unit} gives a real
input for every path, the node sequence it follows, and its expected output;
together they exercise every branch and so achieve path coverage (which subsumes
branch and statement coverage).

\begin{table}[htbp]
\centering
\caption{Path coverage of the air-conditioner target guard ($16$--$30\,$\textdegree
C) with real inputs.}
\label{tab:cfg-unit}
\begin{tabularx}{\textwidth}{@{}l l Y@{}}
\toprule
\textbf{Input $v$} & \textbf{Path (nodes)} & \textbf{Expected output} \\
\midrule
\texttt{"hot"} (non-numeric) & $1\rightarrow2\rightarrow3$ & reject --- not a number \\
$15$ & $1\rightarrow2\rightarrow4\rightarrow5$ & reject --- below 16 \\
$31$ & $1\rightarrow2\rightarrow4\rightarrow6\rightarrow7$ & reject --- above 30 \\
$23$ & $1\rightarrow2\rightarrow4\rightarrow6\rightarrow8$ & accept \\
\bottomrule
\end{tabularx}
\end{table}

The boundary values $16$ and $30$ follow the accepting path
$1\rightarrow2\rightarrow4\rightarrow6\rightarrow8$, $15$ and $31$ are their
just-outside neighbours on the two rejecting branches, and a non-numeric value
(\texttt{"hot"}) takes the remaining path $1\rightarrow2\rightarrow3$; these are
exactly the inputs of UT-1 in Table~\ref{tab:unit-cases}, so that one parametrised
test achieves path coverage of the guard. Table~\ref{tab:unit-cases} lists the full
unit suite.

\begin{table}[htbp]
\centering
\caption{Unit-level test cases (suite \texttt{unit\_domain}); white-box
control-flow coverage. All pass.}
\label{tab:unit-cases}
{\footnotesize
\begin{tabularx}{\textwidth}{@{}L{0.215\textwidth} X X c L{0.085\textwidth}@{}}
\toprule
\textbf{Test case} & \textbf{Input} & \textbf{Expected output} & \textbf{Status} & \textbf{Technique} \\
\midrule
UT-1 \texttt{ac\_target\_boundaries} & AC target $=16,30,23$ / $15,31,$ \texttt{"hot"} & accept / reject & Pass & Path cov. \\
UT-2 \texttt{validate\_rejects\_unknown\_control} & plug, control ``brightness''$=50$ & reject (``no control'') & Pass & Branch cov. \\
UT-3 \texttt{validate\_enum\_and\_toggle} & fan mode \texttt{sleep}/\texttt{turbo}; plug power \texttt{on}/\texttt{maybe} & accept / reject & Pass & Branch cov. \\
UT-4 \texttt{sensor\_is\_readonly} & capability of a sensor & no controls (empty) & Pass & Stmt cov. \\
UT-5 \texttt{flat\_tariff\_pricing} & flat 2000 VND/kWh, 10 kWh & 20\,000 VND & Pass & Stmt cov. \\
UT-6 \texttt{tiered\_tariff\_is\_progressive} & tiers 50@1000 / rest@3000, 100 kWh & 200\,000 VND; marginal@120 $=3000$ & Pass & Path cov. \\
UT-7 \texttt{billing\_cycle\_contains\_now} & ts 2026-06-15, billing day 1 & cycle Jun 1 $\le$ ts $<$ Jul 1 & Pass & Branch cov. \\
UT-8 \texttt{hour\_set\_wraps\_midnight} & window 22:00--04:00 & $\{22,23,0,1,2,3\}$ & Pass & Branch cov. \\
UT-9 \texttt{simulator\_ac\_savings\_differential} & AC at 24\,\textdegree C vs 26\,\textdegree C, 1 day & kWh(24) $>$ kWh(26) $> 0$ & Pass & Path / data-flow \\
UT-10 \texttt{password\_hash\_roundtrip} & hash(``s3cret-pw'') & hash $\ne$ plaintext; verify ok/$\times$ & Pass & Branch cov. \\
\bottomrule
\end{tabularx}}
\end{table}

\section{Integration Testing (White-box)}
\label{sec:qa-integration}

\paragraph{Objective.} Verify that the independently unit-tested components
cooperate correctly once wired together --- that control passes correctly across
the boundaries between router, service, repository, persistence, the in-process
event bus, and the device adapters.

\paragraph{Scope.} The wired request path driven on a fully seeded database
through the Developer's maintenance unit (which owns the complete device package).
The object under test is the \emph{integration logic}: the branches one component
takes depending on another component's state.

\paragraph{Approach.} \emph{Branch / decision coverage of the integrated control
flow}. Because the team knows the internal wiring (white-box), each integration
test is designed to drive a specific cross-component decision to both of its
outcomes --- online vs.\ offline, in-range vs.\ out-of-range, auto-apply on vs.\
off, conflict vs.\ none, undo-within vs.\ after the window.

As the worked example, consider the integrated \emph{command-outcome} path a
device command follows through the service, in the order the code evaluates it: it
\textsc{times out} if the target device is offline; otherwise it is
\textsc{rejected} if the value is out of range; otherwise, if a safety guard
suppresses it --- a safety-critical auto power-off --- it is \textsc{skipped};
otherwise it is applied and \textsc{succeeds}. This path crosses several
components --- the device repository (online state), the capability validator and
the safety guard --- so its CFG (Figure~\ref{fig:cfg-int}) is an \emph{integration}
artefact, with three deterministic decisions and four outcomes.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    node distance=8mm and 16mm,
    block/.style={draw, rounded corners, align=center, font=\scriptsize,
                  minimum width=22mm, minimum height=7mm},
    dec/.style={draw, diamond, aspect=2, align=center, font=\scriptsize,
                inner sep=1pt},
    term/.style={draw, ellipse, align=center, font=\scriptsize, fill=red!8},
    ok/.style={draw, ellipse, align=center, font=\scriptsize, fill=green!12},
    edge/.style={-{Latex[length=2mm]}}
]
\node[block] (m1) {1: command\\$(d, c, v)$};
\node[dec]   (m2) [below=of m1] {2: device\\online?};
\node[term]  (m3) [right=of m2] {3: \textsc{timeout}};
\node[dec]   (m4) [below=of m2] {4: value\\in range?};
\node[term]  (m5) [right=of m4] {5: \textsc{rejected}};
\node[dec]   (m6) [below=of m4] {6: safety\\blocked?};
\node[term]  (m7) [right=of m6] {7: \textsc{skipped}};
\node[ok]    (m8) [below=of m6] {8: \textsc{success}};

\draw[edge] (m1) -- (m2);
\draw[edge] (m2) -- node[above,font=\tiny]{no}  (m3);
\draw[edge] (m2) -- node[right,font=\tiny]{yes} (m4);
\draw[edge] (m4) -- node[above,font=\tiny]{no}  (m5);
\draw[edge] (m4) -- node[right,font=\tiny]{yes} (m6);
\draw[edge] (m6) -- node[above,font=\tiny]{yes} (m7);
\draw[edge] (m6) -- node[right,font=\tiny]{no}  (m8);
\end{tikzpicture}
\caption{Control flow graph of the integrated device-command outcome, in code-evaluation
order. Three decisions (device online, value in range, safety blocked) yield four
outcomes; the graph crosses the device repository, the capability validator and the
safety guard.}
\label{fig:cfg-int}
\end{figure}

Table~\ref{tab:cfg-int} gives a real command for every path.

\begin{table}[htbp]
\centering
\caption{Path coverage of the integrated command-outcome flow with real inputs.}
\label{tab:cfg-int}
\begin{tabularx}{\textwidth}{@{}L{0.36\textwidth} l Y@{}}
\toprule
\textbf{Input (command)} & \textbf{Path (nodes)} & \textbf{Expected outcome} \\
\midrule
offline plug, \texttt{power}$=$off & $1\rightarrow2\rightarrow3$ & \textsc{timeout} \\
AC \texttt{target}$=99$ (out of range), online & $1\rightarrow2\rightarrow4\rightarrow5$ & \textsc{rejected} \\
fridge (safety-critical), \texttt{power}$=$off, online & $1\rightarrow2\rightarrow4\rightarrow6\rightarrow7$ & \textsc{skipped} \\
AC \texttt{target}$=23$, online & $1\rightarrow2\rightarrow4\rightarrow6\rightarrow8$ & \textsc{success} \\
\bottomrule
\end{tabularx}
\end{table}

These four cases are IT-3, IT-2, IT-13 and IT-4 of Table~\ref{tab:int-cases}, which
lists the full integration suite.

\begin{table}[htbp]
\centering
\caption{Integration-level test cases (suites \texttt{devices\_capability},
\texttt{rules}, \texttt{safety}); white-box branch / decision coverage. All pass.}
\label{tab:int-cases}
{\footnotesize
\begin{tabularx}{\textwidth}{@{}L{0.225\textwidth} X X c L{0.085\textwidth}@{}}
\toprule
\textbf{Test case} & \textbf{Input} & \textbf{Expected output} & \textbf{Status} & \textbf{Technique} \\
\midrule
IT-1 \texttt{capability\_endpoint\_drives\_controls} & GET a device's capability & controls $=$ schema set & Pass & Stmt cov. \\
IT-2 \texttt{command\_rejects\_out\_of\_range} & POST AC target $=99$ & 422 rejected & Pass & Branch cov. \\
IT-3 \texttt{offline\_device\_command\_times\_out} & force device offline, POST command & outcome $=$ timeout & Pass & Decision cov. \\
IT-4 \texttt{power\_on\_off\_control} (plug/bulb/fan/AC) & POST power \texttt{on}, then \texttt{off} & state on $\rightarrow$ off & Pass & Branch cov. \\
IT-5 \texttt{add\_and\_delete\_mock\_device} & POST a device, then DELETE it & created (201), then gone & Pass & Path cov. \\
IT-6 \texttt{action\_must\_match\_capability} & rule: plug brightness$=50$ & invalid (422) & Pass & Branch cov. \\
IT-7 \texttt{auto\_action\_off\_by\_default} & create rule, no \texttt{auto\_apply} & \texttt{auto\_apply} $=$ false & Pass & Branch cov. \\
IT-8 \texttt{conflict\_detection} & two rules, same fan, overlapping, on vs.\ off & conflict flagged & Pass & Decision cov. \\
IT-9 \texttt{enable\_disable\_edit\_delete} & create, patch enabled/name, delete & updated, then removed & Pass & Path cov. \\
IT-10 \texttt{time\_rules\_fire\_in\_local\_tz} & rule ``at 19:00'', clock 12:00 UTC & fires ($=$19:00 local); not at 19:00 UTC & Pass & Branch cov. \\
IT-11 \texttt{execution\_logging\_undo\_history} & auto-rule fires, undo, delete rule & success logged; undone; history kept & Pass & Path cov. \\
IT-12 \texttt{ac\_compressor\_rate\_limit} & two compressor commands within 3 min & second rejected & Pass & Decision cov. \\
IT-13 \texttt{safety\_critical\_not\_auto\_off} & auto-rule powers off the fridge & skipped (NFR-SAF-2) & Pass & Decision cov. \\
IT-14 \texttt{safety\_warnings\_on\_validation} & validate temp / safety-critical / unattended-off rule & warnings returned & Pass & Decision cov. \\
IT-15 \texttt{safety\_critical\_not\_forced\_offline} & set the fridge offline & rejected (safety violation) & Pass & Branch cov. \\
IT-16 \texttt{rule\_idempotent\_when\_satisfied} & auto-rule sets AC \texttt{target}$=26$ with the device already at $26$ (stored $26.0$) & no execution and no notification (engine returns \textsc{none}) & Pass & Branch cov. \\
\bottomrule
\end{tabularx}}
\end{table}

IT-16 and IT-11 together exercise both outcomes of the rule engine's
\emph{idempotency} decision, which precedes command dispatch: IT-11 drives the case
where the device is not yet in the rule's target state (the rule fires and is logged),
while IT-16 drives the complementary case where the device already holds the target
value (the rule is suppressed, producing no execution and no notification). The IT-16
input is chosen with white-box knowledge that a \textsc{range} value is stored in
normalized form (target $26$ as $26.0$), so the test guards against a
literal-versus-normalized mismatch silently re-firing the rule.

\section{System Testing (Black-box)}
\label{sec:qa-system}

\paragraph{Objective.} Validate the user-visible behaviour of the whole system
against its functional specification, treating the server as a black box reached
only through its public REST interface --- without reference to internal structure.

\paragraph{Scope.} The fully seeded application driven end-to-end through its
endpoints: monitoring and reporting, bill-saving, recommendations, authentication
and role-based access control, and the Administrator's building oversight.

\paragraph{Approach.} The four standard black-box test-case design techniques,
applied per case rather than ad hoc: \emph{equivalence partitioning} (EP, one
representative per valid/invalid input or output class --- e.g.\ the caller-role
classes and credential classes), \emph{boundary-value analysis} (BVA, e.g.\ the
``at most five recommendations'' cap), \emph{decision tables} (an outcome from a
combination of conditions, e.g.\ the saving-estimate formula), and
\emph{state-based} testing (lifecycle transitions, e.g.\ a recommendation moving
from active to accepted or dismissed, and a unit from occupied to vacant). The
technique used for each case is given in the last column of
Table~\ref{tab:sys-cases}.

\begin{table}[htbp]
\centering
\caption{System-level test cases (suites \texttt{auth\_rbac}, \texttt{admin},
\texttt{monitoring}, \texttt{savings}, \texttt{recommendations}); black-box. All pass.}
\label{tab:sys-cases}
{\footnotesize
\begin{tabularx}{\textwidth}{@{}L{0.225\textwidth} X X c L{0.06\textwidth}@{}}
\toprule
\textbf{Test case} & \textbf{Input} & \textbf{Expected output} & \textbf{Status} & \textbf{Technique} \\
\midrule
ST-1 \texttt{login\_returns\_token\_and\_user} & valid credentials & 200 + token + user & Pass & EP \\
ST-2 \texttt{login\_rejects\_bad\_password} & wrong password & 401 & Pass & EP \\
ST-3 \texttt{unauthenticated\_is\_rejected} & request with no token & 401 & Pass & EP \\
ST-4 \texttt{me\_returns\_current\_user} & GET \texttt{/me} with valid token & 200, the caller & Pass & EP \\
ST-5 \texttt{resident\_cannot\_add\_device} & Resident POST device & 403 & Pass & EP \\
ST-6 \texttt{resident\_cannot\_create\_tariff} & Resident POST tariff & 403 & Pass & EP \\
ST-7 \texttt{admin\_cannot\_control\_device} & Administrator POST command & 403 & Pass & EP \\
ST-8 \texttt{admin\_cannot\_add\_device} & Administrator POST device & 403 & Pass & EP \\
ST-9 \texttt{resident\_cannot\_access\_overview} & Resident GET building overview & 403 & Pass & EP \\
ST-10 \texttt{resident\_cannot\_offboard} & Resident DELETE resident & 403 & Pass & EP \\
ST-11 \texttt{admin\_building\_overview} & Administrator GET overview & per-unit aggregates & Pass & EP \\
ST-12 \texttt{admin\_can\_drill\_into\_a\_unit} & Administrator GET a unit & that unit's detail & Pass & EP \\
ST-13 \texttt{admin\_has\_no\_unit} & Administrator GET \texttt{/me} & \texttt{home\_id} $=$ null & Pass & EP \\
ST-14 \texttt{admin\_can\_onboard\_resident} & Administrator onboards a vacant unit & created; resident scoped to own unit & Pass & State \\
ST-15 \texttt{admin\_offboard\_is\_soft} & Administrator DELETE resident & 204; cannot re-authenticate; unit vacant & Pass & State \\
ST-16 \texttt{offboard\_vacant\_is\_conflict} & offboard an already-vacant unit & 409 conflict & Pass & State \\
ST-17 \texttt{dashboard\_reports\_metrics} & GET dashboard & live power + kWh + bill present & Pass & EP \\
ST-18 \texttt{top\_consumers\_ranked} & GET top consumers & descending by energy & Pass & EP \\
ST-19 \texttt{consumption\_series\_buckets} & GET series, granularity hour & series bucketed by hour & Pass & EP \\
ST-20 \texttt{consumption\_rejects\_foreign} & Resident GET another unit's device & denied (NFR-SEC-4) & Pass & EP \\
ST-21 \texttt{csv\_export} & GET CSV export & well-formed CSV & Pass & EP \\
ST-22 \texttt{estimate\_uses\_srs\_formula} & estimate a rule before saving & equals SRS formula; baseline $>0$ & Pass & Decision \\
ST-23 \texttt{raising\_ac\_target\_saves} & rule raising the AC target & positive saving & Pass & EP \\
ST-24 \texttt{savings\_summary\_reports\_cycle} & GET savings summary & current billing-cycle savings & Pass & EP \\
ST-25 \texttt{savings\_records\_present} & GET savings records & records exist & Pass & EP \\
ST-26 \texttt{drift\_check\_is\_reachable} & run the drift check & drift flag returned & Pass & EP \\
ST-27 \texttt{recommendations\_readable} & analyse, then GET recommendations & WHEN/THEN + rationale + window & Pass & EP \\
ST-28 \texttt{recommendations\_ranked\_capped} & analyse usage & sorted by VND, at most five & Pass & BVA \\
ST-29 \texttt{new\_device\_yields\_no\_rec} & add a fresh device, analyse & none for it ($<7$ days data) & Pass & EP \\
ST-30 \texttt{recommendation\_provider\_swappable} & inject a fake provider & default Heuristic; injected one consulted & Pass & Seam \\
ST-31 \texttt{accept\_becomes\_rule} & POST accept a recommendation & 201 rule; no longer active & Pass & State \\
ST-32 \texttt{dismiss\_suppresses} & POST dismiss, then re-analyse & suppressed (not re-shown) & Pass & State \\
\bottomrule
\end{tabularx}}
\end{table}

\section{Results and Traceability}
\label{sec:qa-results}

The full suite passes: \textbf{67 tests, 67 passed}, as recorded in the
\emph{Status} column of every table above. Because tests are derived from the
specification, every requirement family is traceable to at least one executable
check, and conversely every test traces back to a requirement.
\emph{Acceptance testing} re-runs this same system suite against the demonstration
seed in front of the customer: the green system suite \emph{is} the evidence that
the acceptance focus items hold. Table~\ref{tab:trace} summarises the mapping from
requirement families to representative covering tests.

\begin{table}[htbp]
\centering
\caption{Requirements-to-tests traceability (representative tests).}
\label{tab:trace}
\begin{tabularx}{\textwidth}{@{}l L{0.32\textwidth} Y@{}}
\toprule
\textbf{Requirement family} & \textbf{Concern} & \textbf{Behaviour verified} \\
\midrule
REQ-4.1 & Monitoring (live power, cumulative energy, offline detection) &
  ST-17 dashboard metrics; IT-3 command to an offline device times out \\
REQ-4.2 & Device control via capability schema (load, validate, command) &
  IT-1 the schema drives the controls; IT-2 an out-of-range command is rejected \\
REQ-4.3 & Rules, conflict detection, auto-action opt-in and undo &
  IT-8 conflicting rules detected; IT-7 automatic actions off by default \\
REQ-4.4 & Recommendation generation, ranking and approval &
  ST-31 an accepted recommendation becomes a rule; ST-28 ranked and capped \\
REQ-4.5 & Bill-saving estimation and cycle reporting &
  ST-22 estimate follows the formula; ST-24 summary reports the cycle \\
NFR-SAF & Compressor rate limit, safety-critical protection, warnings &
  IT-12 compressor rate limit; IT-13 safety-critical auto-power-off suppression \\
NFR-SEC & Authentication, RBAC, no plaintext secrets &
  ST-3 unauthenticated request rejected; UT-10 password hash round-trip \\
\bottomrule
\end{tabularx}
\end{table}

Both styles of testing are present and traceable: white-box unit and integration
tests pin the algorithms and the integrated control flow, while black-box system
tests driven through the public interface prove the wired system behaves
correctly. Negative and boundary paths are first-class throughout: rejected,
timeout and skipped command outcomes, out-of-range and unsupported controls, bad
passwords, unauthenticated access and forbidden cross-role actions are all
explicitly exercised. One boundary is asserted only on its inner side: ST-28
confirms the recommendation list never exceeds the cap of five, but the
demonstration seed yields fewer than five candidates, so the exact transition (a
sixth recommendation being trimmed) is not forced. The principal residual gaps ---
that boundary, plus direct assertion of notification delivery, frontend rendering,
and non-functional load and transport security --- are documented as known
limitations to be closed with a six-candidate cap test, a notification endpoint
test, manual UI acceptance, and a dedicated load and TLS harness respectively.

\section{Maintenance}
\label{sec:qa-maintenance}

Verification and validation do not end at delivery; they continue through the
Support phase of the life-cycle. Anticipated changes to SHEO classify into the
four standard categories of software maintenance, each of which would re-enter the
regression suite before release:

\begin{itemize}[leftmargin=1.4em]
  \item \textbf{Corrective} --- repairing a latent fault. For example, fixing an
        off-by-one error should the air-conditioner target guard ever admit $31$;
        the boundary case UT-1 in Table~\ref{tab:unit-cases} would reproduce and
        then guard the fix.
  \item \textbf{Adaptive} --- accommodating a changed environment. For example,
        updating the tariff math when the electricity utility revises its tiered
        price schedule, re-validated by the tariff-pricing unit tests (UT-5, UT-6).
  \item \textbf{Perfective} --- improving function or quality without changing
        external behaviour. For example, adding a new device type (a water
        heater) by declaring one capability schema entry and one adapter, then
        extending the capability and command tests to cover it.
  \item \textbf{Preventive} --- restructuring to forestall future defects. For
        example, hardening the rate-limiter against clock skew before any failure
        is observed, protected by the existing rate-limit regression test (IT-12).
\end{itemize}

This classification realises the Support phase of the life-cycle introduced in
the process chapter, and closes the verification-and-validation loop: each kind
of change is admitted only once the relevant tests --- old and new --- pass,
keeping the requirements-to-tests traceability of Table~\ref{tab:trace} honest as
the system evolves.
