\begin{thebibliography}{99}
\addcontentsline{toc}{chapter}{References}

\bibitem{c01-09} Introduction to Software Engineering Teaching Group, \emph{IT3180E --- Introduction to Software Engineering}, Lecture slides C01--C09 (Overview of SE; Software Development Process; Agile Software Development; Software Project Management; Software Configuration Management; Requirement Analysis; Software Design and User Interface Design; Software Construction; Quality Assurance). Hanoi University of Science and Technology, SoICT, 2025.

\bibitem{ieee1016} IEEE, \emph{IEEE Std 1016 --- IEEE Standard for Information Technology --- Systems Design --- Software Design Descriptions}. Institute of Electrical and Electronics Engineers.

\bibitem{ieee29148} ISO/IEC/IEEE, \emph{ISO/IEC/IEEE 29148 --- Systems and Software Engineering --- Life Cycle Processes --- Requirements Engineering} (superseding IEEE Std 830).

\bibitem{ieee828} IEEE, \emph{IEEE Std 828 --- IEEE Standard for Configuration Management in Systems and Software Engineering}.

\bibitem{ieee610} IEEE, \emph{IEEE Std 610.12-1990 --- IEEE Standard Glossary of Software Engineering Terminology}.

\bibitem{iso9126} ISO/IEC, \emph{ISO/IEC 9126 --- Software Engineering --- Product Quality}.

\bibitem{pressman} R.~S.~Pressman and B.~R.~Maxim, \emph{Software Engineering: A Practitioner's Approach}, 8th ed. McGraw-Hill.

\bibitem{sommerville} I.~Sommerville, \emph{Software Engineering}, 10th ed. Pearson.

\bibitem{gof} E.~Gamma, R.~Helm, R.~Johnson, and J.~Vlissides, \emph{Design Patterns: Elements of Reusable Object-Oriented Software}. Addison-Wesley.

\bibitem{fowler} M.~Fowler, \emph{Refactoring: Improving the Design of Existing Code}, 2nd ed. Addison-Wesley.

\bibitem{agile} K.~Beck et al., \emph{Manifesto for Agile Software Development}, 2001. \url{https://agilemanifesto.org}

\bibitem{fastapi} S.~Ram\'irez, \emph{FastAPI Documentation}. \url{https://fastapi.tiangolo.com}

\bibitem{sqlalchemy} \emph{SQLAlchemy 2.0 Documentation}. \url{https://docs.sqlalchemy.org}

\bibitem{react} \emph{React Documentation}. \url{https://react.dev}

\end{thebibliography}
